Typed Functional Genetic Programming

### 2013-07-09 tex/thesis.tex
[tex]

@@

@@

Therefore $building$ $blocks = T \cup F$.\\

Beside symbols in $T \cup F$ there must also be 
an implementation for each symbol which is not variable. 
And for every function symbol from $F$ there must be specified 
number of arguments.\\

@@

Satisfying this constrain ensures that every program tree build 
from $T \cup F$ will be total.\\

@@

We will describe tree generating method used by Koza in \cite{koza92} called
\textit{ramped half-and-half}.

\textbf{TODO!}\\
\textbf{Talk about Generating initial population}

@@

@@

Selection of genetic operation in each operation is 
controlled by probabilities for each genetic operator.

@@

\begin{itemize}
	\item \textit{Crossover}    : $70\%$
	\item \textit{Reproduction} : $30\%$
	\item \textit{Mutation}     :  $0\%$
\end{itemize}

@@

Bellow is described GP algorithm in pseudocode.
Inputs are:

\begin{itemize}
	\item \textit{fitness} - Fitness function.
	\item \textit{TuF} - Building blocks accompanied with an  	
	      information about implementations and argument numbers.
	\item \textit{popSize} - Population size.
	\item \textit{numGens} - Number of generations.
	\item \textit{probabs} - Genetic operators probabilities.
\end{itemize}

\Pseudokod{GP(fitness,TuF,popSize,numGens,probabs)}{
	$gen \leftarrow$ 0 \;
	$pop \leftarrow generateInitialPopuletion( TuF ) $ \;
	($popWithF,terminate,best) \leftarrow$ 
	evaluate($fitness$, $TuF$, $pop$)\;

	\While{ $gen < numGens$ $\wedge$ $\neg terminate$  }{

	$newPop \leftarrow$ empty population \;
	$newPop$.insert( $best$ )\;
	$i \leftarrow 1$ \;

	\While{ $i < popSize$ }{
		$op \leftarrow probabilisticallySelectOperation(probabs)$ \;		
		\Switch{op}{
			\Case{Crossover}{
				$parent1 \leftarrow$ selection( $popWithF$ ) \;
				$parent2 \leftarrow$ selection( $popWithF$ ) \;
				($child1$,$child2$) = crossover( $parent1$ , $parent2$ )\;
				$newPop$.insert( $child1$ ) \;
				$newPop$.insert( $child2$ ) \;
				$i \leftarrow i + 2$ \;
			}
			\Case{Reproduction}{
				$indiv \leftarrow$ selection( $popWithF$ ) \;
				$newPop$.insert( $indiv$ ) \;
				$i \leftarrow i + 1$ \;
			}
			\Case{Mutation}{
				$indiv \leftarrow$ selection( $popWithF$ ) \;
				$mutant \leftarrow$ mutate( $indiv$ , $TuF$ ) \;
				$newPop$.insert( $mutant$ ) \;
				$i \leftarrow i + 1$ \;			
			}		
		}
	}

	$pop \leftarrow newPop$  \;
	($popWithF,terminate,best) \leftarrow$ 
	evaluate($fitness$, $TuF$, $pop$)\;
	$gen \leftarrow gen + 1$ \; 
 }
 \Return pop \;
}

In order to clarify the code let us describe in greater detail inputs and
contained procedures
generateInitialPopuletion(),
evaluate(),
probabilisticallySelectOperation(),
selection(),
crossover() 
and mutation().

\textbf{TODO mention best preservation and same popSize checking. }

...\\\
\subsection{TODO}
	\begin{itemize} 
		\item TALK ABOUT GP is part of EA etc. and maybe define the GP by defining EA and then
		      specifying the differences or something like that...    
		\item History, citations, etc ....
		\item ten algoritmus v kozovi dělá "Designate Result" já tam vracim poslední populaci
	\end{itemize}

@@

\newcommand{\lamb}[2]{( \lambda #1 . #2 )}
\newcommand{\lam}[2]{\lambda #1 . #2}

Let $V$ be infinite countable set of {\it 
variable names}.  \\* 
Let $C$ be set of {\it constant names}.	 \\*		
Then $\Lambda$ is set of {\it \lterms} defined inductively as follows
\footnote{Function application and $\lambda$-abstraction are concepts
well known from common programming languages. 
For example in JavaScript 
$(M~N)$ translates to expression \texttt{$M$($N$)} and
$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
In other words, the $\lambda$-abstraction is equivalent to 
\textit{anonymous function}. 
}.

\begin{align*}
x   \in V \cup C  &\then x     \in \Lambda \\
M,N \in \Lambda   &\then (M~N) \in \Lambda 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
\textit{~~~~~($\lambda$-abstraction)} 
\end{align*}
\\
It is usual to use $V = \Sigma^+$ set of all non-empty finite strings of symbols 
from $\Sigma$ where $\Sigma$ is some alphabet set, e.g.  
$
\Sigma =
\setDots{a}{z} \cup 
\setDots{_0}{_9} $.

@@

We use upper case symbols such as 
$M,N,M_1,M_2,...,N_1,N_2,...$
to denote arbitrary lambda terms in contrast with
those in lower case such as
$a,b,c,...,x,y,z,$
$f_1,f_2,...,g_1,g_2,...$
denoting variable names, elements of $V$.
Sometimes we will use lower case symbols
to denote arbitrary variable names (such as
$x$ in $\lamb{x}{M}$),
and at other times, we will use them
as specific variable names in specific terms 
(such as $x$ in $\lamb{x}{x}$).
The distinction between the two should be
clear from the context.\\

@@

\begin{easylist}[itemize]
& $x$
& $\lamb{x}{x}$
& $(\lamb{x}{(x~x)}~\lamb{x}{(x~x)})$
& $\lamb{x}{\lamb{y}{x}}$
& $\lamb{f_1}{\lamb{f_2}{\lamb{x}{((f_1~x)~(f_2~x))}}}$
\end{easylist}~

@@

For example we can write
$\lamb{f_1}{\lamb{f_2}{\lamb{x}{((f_1~x)~(f_2~x))}}}$
as $\lam{f_1 f_2 x}{f_1~x~(f_2~x)}$.

@@

@@

\textbf{TODO možná zmínit že y nesmí být v N volně v připadě\\
$\lamb{y}{P}[x:=N] = \lam{y}{( P[x:=N] )}$ }, ale řekl bych že zbytečnej detail

@@

@@

~\\\\
\textbf{TODO} 
\begin{itemize} 
	\item promluvit krátce o intuici
	      a nějak to napojit na to jak to používám já
	      pak v tom GP.
	\item Citovat nějakýho bárendrechta jako zdroj podrobnějších informací
\end{itemize}

@@

~\\
\textbf{TODO} 
\begin{itemize}
	\item určitě promluvit nějak v úvodu k tomu že se jedná o 
	Simply typed lambda kalkulus, zmínit že klasicky se těm alfa z A 
	řiká proměný, ale že je to matoucí vzhledem k polymorfizmu
	\item možná se hlobějc zmínit o curryingu  
\end{itemize}

@@

	In other words context is a set of statements with distinct variables or constants as subjects.
	\\\\
	\textbf{TODO:} TALK ABOUT Context represents library/building blocks.

@@

\textbf{TODO : popsat tu jak ji předelat na beta nf
pokud by se nám chtělo... 
V barendrech84 (ta nová kniha s žlutejma deskama) je ukázano že je 
eta CR (čurč roserovská) což snad stačí společně s 
tim tvrzeničkem z book.pdf že vezmu lnf
a dělám eta redukce dokud to de. až to nejde tak bych to měl mít v 
beta-nf }

~\\\textbf{TODO : přepsat to nějak hezčejc, ten seznam heskejch vlastností}

@@

Inference rule 2: 
\[
	\frac{ (f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma \qquad
	       \Gamma \vdash M_1 : \rho_1 \quad
	       \dotsm \quad
	       \Gamma \vdash M_m : \rho_m        
	      }
	     {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
\]
\\
Proof of correctness (\textbf{TODO REPAIR} Conceptually it is ok but there is sazba-bug somewhere): 
\[
   \dfrac
     {\dfrac
      {\dfrac
       {\dfrac         
         {\dfrac  
          {\dfrac
           {\boxed{(f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma}}
           {\Gamma \vdash f : \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha}
           \quad
           \boxed{\Gamma \vdash M_1 : \rho_1} }
          {\Gamma \vdash (f M_1) : \rho_2 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha }
          }{\vdots} 
         \quad 
         \ddots }
       {\Gamma \vdash (f M_1 \dots M_{m-2}) : \rho_{m-1} \rightarrow \rho_m \rightarrow \alpha}
       \quad
       \boxed{\Gamma \vdash M_{m-1} : \rho_{m-1}}  }
      {\Gamma \vdash (f M_1 \dots M_{m-1}) : \rho_m \rightarrow \alpha}       
      \quad 
      \boxed{\Gamma \vdash M_m : \rho_m} }
	 {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
\]
\\
... there is a grammar rule:
\[ 
	( \alpha , \Gamma )  \longmapsto
	\bigg( \mbox{ {\Large f }}
	  \mbox{ \Vtextvisiblespace[1em] } 
	  ( \rho_1 , \Gamma )
	  \mbox{ \Vtextvisiblespace[1em] } 
	  \dots
	  \mbox{ \Vtextvisiblespace[1em] } 
	  ( \rho_m , \Gamma )
	  \quad \bigg)
\]
\\\\
\textbf{TODO} 
\begin{itemize}
	\item SHOW correctness of those inference rules by composing them of 
		  $E^{\rightarrow}$, $I^{\rightarrow}$ and \textit{axiom}.
	\item SHOW more examples of inference rules transformed into grammar rules.
	\item DESCRIBE general algorithm for this transformation.
	\item TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
	\item TALK ABOUT $\beta \eta^{-1}$-normal form which is generated by this method.
\end{itemize}

~\\\textbf{TODO : Napsat tu (nebo asi někam jinam
ale hlavně někam), že de generovat přímo i termy v beta normalní 
formě, ale že se mi to zdálo o dost pracnější díky tomu, že
musíme čekovat jestli neni něco toho typu i v ne atomickejch
vrcholech/typech... }

@@

~\\
\textbf{TODO} 
\begin{itemize}
\item EXAMPLES of tree representations of \lterms  
\item TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
		Things about $\beta\eta^{-1}$ normal form, etc.   
\item DESCRIBE Inhabitation Machine... 
\end{itemize}

@@

\textbf{TODO\\
Dřív se to menovalo 
"Our approach to \textit{solving} Inhabitation tree"
tak tu napsat pěknej uvod o tom, že naši jedinci jsou lambda
termy a že je  budeme generovat pomocí 
řešení inhabitačních stromů.\\\\
}

@@

\textit{Classical-GP-like} ... \textbf{TODO}

@@

\textbf{TODO : víc to tu rozepsat hlavně zmínit heuristiku 
a že stačí aby byla optimistická atd.. }

@@

...\\\\\\\\
\textbf{TODO : čeknout jak je to v kódu jestli ten A* co tam je čekuje ty
 nódy co už navštívil při přidávání - VŽDYŤ díky tomu žeje to tree-search 
 (vs. graph search) to neni potřeba - pořádně se podívat do AIAMA jakj
 je terminologie a napsat to sem.
}

@@

\textbf{TODO : Discuss that it is OK to expand only the DFS-found node. 
The order of unfinished nodes is irrelevant since generation of each branch is 
independent on generating other branches. Můžem na to koukat jako na takovou tu redukci 
simetrií, najít to v AIAMA a pořádně popsat.  }

\textbf{TODO : Odiskutovat nějak tohle:
přirozeně chceme při produkci velkýho množství termů obhospodařovávat nějaký
velký množství rozdělanejch termů v nějaký struktuře a postuipně je dodělávat.
Proto přichází jako tahle struktura na mysl nějaká prioritní fronta,
otázka je kde vzít ty priority a A* je jednoduchá ale přitom optimum hledající 
odpověď na tuhle otázku, dostatečně volná skrz tu heuristiku.  }

\textbf{TODO: určitě někde zmínit ospravedlnění proč si napsat dokazovač sám
a né požít nějakej už hotovej - de o to, že dokazovače jsou delaný na to aby
udělali jeden důkaz typicky složitýho tvrzení, my sme v situaci kdy
chceme co nejvíc (a co možná co nejvíc rozmanitejch) důkazů jednoduchýho problému.
 }

@@

\textbf{TODO : rozebrat to, ukázat že je optimistická atd...}

@@

\textbf{TODO}
\begin{enum}
& Prostě budem křížit jen podstromy stejnýho typu
& ALE! Hlavní problém jsou proměnný
& ve stručnosti kouknout na dvě hlavní metody
  && křížit prostě lambda stromy a vzniklý problémy řešit online 
  && zbavit se proměnejch 
& říct že nakonec sme přilnuli k zbavit se proměnejch
\end{enum}

@@

\subsection{Conversion to SKI combinators}
\textbf{TODO}

@@

\textbf{TODO: 
říct že na tu sme se moc nesoustředili ze dvou důvodů
ktery to ospravedlňujou \\
1) koza taky\\
2) ta s vygereovanim podstromu je v zásadě použití
   algoritmu na generovaní jedince a tak nám přišlo
   lepší pořádně se soustředit na generování jedince a
   začít do toho mutaci tahat až bude generování pořádně
   prozkoumáno.}

@@

>>>>

% !!! místo we can atd psát one can ...

@@

\newenvironment{todo}
{ ~\\[1em]
  \textbf{TODO}
  \begin{easylist}[itemize]}
{\end{easylist}~\\[1em]}

@@

\newcommand{\TuF}{$T \cup F$\xspace}

Therefore $building$ $blocks = $ \TuF.\\

Beside symbols in \TuF there must also be 
an implementation for each symbol which is not variable. 
And for every function symbol from $F$ there must be specified 
number of arguments.\\

@@

Satisfying this constrain ensures that every program tree build 
from \TuF will be total.\\

@@

We will describe three tree generating method described by Koza 
in \cite{koza92} called \textit{ramped half-and-half}. \\

An individual tree is constructed by random (uniform) selection of symbol from 
subset of \TuF for the root node and after that by generation of its 
subtrees recursively. 
Number of subtrees of a node corresponds 
to the number of arguments for the selected symbol.

There are two generating sub-methods called \textit{full} and 
\textit{grow} each restricting differently the subset of \TuF. 
This restriction depends on depth of the
node for which the symbol is being selected.

In order to perform \textit{ramped half-and-half} method
one of those two sub-methods is randomly (uniformly) 
selected and the selected one is performed.\\

In the root (\textit{depth} = 0) both \textit{full} and \textit{grow}
select the symbol from the set $F$.

A maximum depth $D_{initial}$ is defined, i.e. 6.

In nodes with \textit{depth} $\leq D_{initial}$ the \textit{full}
method selects from the set $F$, whereas the \textit{grow} method 
selects from whole \TuF.

In nodes with \textit{depth} $= D_{initial}$ both methods select from the
set $T$.

@@

\begin{todo}
 & zmínit maximální hloubku vzniklýho dítěte (17)
\end{todo}

@@

Selection of genetic operation in each operation is 
controlled by probabilities for each genetic operator.\\

@@

\begin{itemize}
	\item \textit{Crossover}    : $90\%$
	\item \textit{Reproduction} : $10\%$
	\item \textit{Mutation}     :  $0\%$
\end{itemize}

@@

Bellow is described GP algorithm in pseudocode.

\Pseudokod{GP( fitness, \TuF, popSize, numGens, probabs )}{
  $gen \leftarrow$ 0 \;\;

  $pop \leftarrow generateInitialPopuletion( $ \TuF$ ) $ \;
  ($popWithF,terminate,best) \leftarrow$ 
  evaluate($fitness$, \TuF, $pop$)\;\;

  \While{ $gen < numGens$ $\wedge$ $\neg terminate$  }{  	
	$newPop \leftarrow$ empty population \;
	$newPop$.insert( $best$ )\;\;

	$i \leftarrow 1$ \;
	\While{ $i < popSize$ }{	
		$op \leftarrow probabilisticallySelectOperation(probabs)$ \;\;

		\Switch{op}{

			\Case{Crossover}{
				$parent1 \leftarrow$ selection( $popWithF$ ) \;
				$parent2 \leftarrow$ selection( $popWithF$ ) \;\;

				($child1$,$child2$) = crossover( $parent1$ , $parent2$ )\;\;

				$newPop$.insert( $child1$ ) \;
				$newPop$.insert( $child2$ ) \;
				$i \leftarrow i + 2$ \;
			}
			\Case{Reproduction}{
				$indiv \leftarrow$ selection( $popWithF$ ) \;
				$newPop$.insert( $indiv$ ) \;
				$i \leftarrow i + 1$ \;
			}
			\Case{Mutation}{
				$indiv \leftarrow$ selection( $popWithF$ ) \;
				$mutant \leftarrow$ mutate( $indiv$ , \TuF ) \;
				$newPop$.insert( $mutant$ ) \;
				$i \leftarrow i + 1$ \;	
			}		
		}
	}\;

	$pop \leftarrow newPop$  \;
	($popWithF,terminate,best) \leftarrow$ 
	evaluate($fitness$, $TuF$, $pop$)\;
	$gen \leftarrow gen + 1$ \; 
 }\;

 \Return pop \;
}

\begin{samepage}
\Lets clarify the input arguments:
\begin{itemize}
	\item \textit{fitness} - Fitness function.
	\item \textit{\TuF} - Building blocks accompanied with an  	
	      information about implementations and argument numbers.
	\item \textit{popSize} - Population size.
	\item \textit{numGens} - Number of generations.
	\item \textit{probabs} - Genetic operators probabilities.
\end{itemize} 
\end{samepage}

%In order to clarify the code let us describe in greater detail inputs and

Behaviors of contained procedures
generateInitialPopuletion(),
evaluate(),
probabilisticallySelectOperation(),
selection(),
crossover() 
and mutation() are hopefully clear from verbal description.

\begin{todo}
 & mention best preservation and same popSize checking.
 & TALK ABOUT GP is part of EA etc. and maybe define the GP by 
   defining EA and then specifying the differences or something 
   like that...    
 & History, citations, etc ....
 & ten algoritmus v kozovi dělá "Designate Result" já tam vracim 
   poslední populaci
\end{todo}

@@

\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}

\newcommand{\Scomb }{\mathbf{S}}
\newcommand{\Kcomb }{\mathbf{K}}
\newcommand{\Icomb }{\mathbf{I}}

Let $V$ be infinite countable set of {\it 
variable names}.  \\* 
Let $C$ be set of {\it constant names}.	 \\*		
Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
\begin{align*}
x   \in V \cup C  &\then x     \in \Lambda \\
M,N \in \Lambda   &\then (M~N) \in \Lambda 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
\textit{~~~~~($\lambda$-abstraction)} 
\end{align*}

Function application and $\lambda$-abstraction are concepts
well known from common programming languages. 
For example in JavaScript 
$(M~N)$ translates to expression \texttt{$M$($N$)} and
$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
In other words, the function application 
corresponds to the act of supplying a function 
with an argument and
the $\lambda$-abstraction is equivalent to 
\textit{anonymous function}. \\

It is usual to use $V = \Sigma^+$ set of all non-empty finite strings of symbols 
from $\Sigma$ where $\Sigma$ is some alphabet set, e.g.  
$
\Sigma =
\setDots{a}{z}$.

@@

We use upper case symbols such as 
$M,N,M_1,M_2,...,N_1,N_2,...$
to denote arbitrary lambda terms in contrast with
those in lower case such as
$a,b,c,...,x,y,z,$
$f_1,f_2,...,g_1,g_2,...$
denoting variable names, elements of $V$.

Sometimes we will use lower case symbols
to denote arbitrary variable names (such as
$x$ in $\lamb{x}{M}$),
and at other times, we will use them
as specific variable names in specific terms 
(such as $x$ in $\lamb{x}{x}$).
The distinction between the two should be
clear from the context.

We use bold upper case symbols such as 
$\Scomb, \Kcomb, \Icomb$ as abbreviation
for specific terms, e.i. 
$\Icomb = \lamb{x}{x}$.  
\\

@@

\begin{easylist}[itemize]
& $x$
& $\lamb{x}{x}$
& $\lamb{x}{f}$
& $(\lamb{x}{(x~x)}~\lamb{x}{(x~x)})$
& $\lamb{x}{\lamb{y}{x}}$
& $\lamb{f}{\lamb{g}{\lamb{x}{((f~x)~(g~x))}}}$
\end{easylist}~

@@

For example we can write
$\lamb{f}{\lamb{g}{\lamb{x}{((f~x)\,(g~x))}}}$
as $\lam{f\,g\,x}{f~x\,(g~x)}$.

@@

A variable in \lterm $M$ is called \textit{bound} if it is not free.

\lterm M is called \textit{combinator} if $\FV(M)=\emptyset$.\\

Notable examples of combinators are combinators
$\Scomb$, $\Kcomb$ and $\Icomb$.
\begin{align*}
\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
\Kcomb &= \lam{x\,y}{x} \\
\Icomb &= \lam{x}{x} 
\end{align*}

As will be shown in \ref{toSKI}
these combinators will help us in performing crossover of \lterms.\\

@@

\begin{todo}
& možná zmínit že y nesmí být v N volně v připadě\\
  $\lamb{y}{P}[x:=N] = \lam{y}{( P[x:=N] )}$ , 
  ale řekl bych že zbytečnej detail
\end{todo}

@@

In other words, $\beta$-reduction is the process 
of insertion of arguments supplied to a function into 
its body.

@@

\begin{todo}
	& říct co to vubec je ty intutivně (v par větách jen) 
	  a nějak to napojit na to jak to používám já
	  pak v tom GP.
	& Citovat nějakýho bárendrechta jako zdroj podrobnějších informací
\end{todo}

@@

\begin{todo}
   & určitě promluvit nějak v úvodu k tomu že se jedná o 
	 Simply typed lambda kalkulus, zmínit že klasicky se těm alfa z A 
	 řiká proměný, ale že je to matoucí vzhledem k polymorfizmu
   & možná se hlobějc zmínit o curryingu  
\end{todo}

@@

	In other words context is a set of statements with distinct variables or constants as subjects.

\begin{todo}
 & TALK ABOUT Context represents library/building blocks.
\end{todo}

@@

\begin{todo}
 & popsat tu jak ji předelat na beta nf
	pokud by se nám chtělo... 
	V barendrech84 (ta nová kniha s žlutejma deskama) je ukázano že je 
	eta CR (čurč roserovská) což snad stačí společně s 
	tim tvrzeničkem z book.pdf že vezmu lnf
	a dělám eta redukce dokud to de. až to nejde tak bych to měl mít v 
	beta-nf
 & přepsat to nějak hezčejc, ten seznam heskejch vlastností
\end{todo}

@@

Inference rule 2: 
\[
	\frac{ (f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma \qquad
	       \Gamma \vdash M_1 : \rho_1 \quad
	       \dotsm \quad
	       \Gamma \vdash M_m : \rho_m        
	      }
	     {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
\]
\\
Proof of correctness (\textbf{TODO REPAIR} Conceptually it is ok but there is sazba-bug somewhere): 
\[
   \dfrac
     {\dfrac
      {\dfrac
       {\dfrac         
         {\dfrac  
          {\dfrac
           {\boxed{(f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma}}
           {\Gamma \vdash f : \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha}
           \quad
           \boxed{\Gamma \vdash M_1 : \rho_1} }
          {\Gamma \vdash (f M_1) : \rho_2 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha }
          }{\vdots} 
         \quad 
         \ddots }
       {\Gamma \vdash (f M_1 \dots M_{m-2}) : \rho_{m-1} \rightarrow \rho_m \rightarrow \alpha}
       \quad
       \boxed{\Gamma \vdash M_{m-1} : \rho_{m-1}}  }
      {\Gamma \vdash (f M_1 \dots M_{m-1}) : \rho_m \rightarrow \alpha}       
      \quad 
      \boxed{\Gamma \vdash M_m : \rho_m} }
	 {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
\]
\\
... there is a grammar rule:
\[ 
	( \alpha , \Gamma )  \longmapsto
	\bigg( \mbox{ {\Large f }}
	  \mbox{ \Vtextvisiblespace[1em] } 
	  ( \rho_1 , \Gamma )
	  \mbox{ \Vtextvisiblespace[1em] } 
	  \dots
	  \mbox{ \Vtextvisiblespace[1em] } 
	  ( \rho_m , \Gamma )
	  \quad \bigg)
\]

\begin{todo}
& SHOW correctness of those inference rules by composing them of 
	  $E^{\rightarrow}$, $I^{\rightarrow}$ and \textit{axiom}.
& SHOW more examples of inference rules transformed into grammar rules.
& DESCRIBE general algorithm for this transformation.
& TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
& TALK ABOUT $\beta \eta^{-1}$-normal form which is generated by this method.
& Napsat tu (nebo asi někam jinam
	ale hlavně někam), že de generovat přímo i termy v beta normalní 
	formě, ale že se mi to zdálo o dost pracnější díky tomu, že
	musíme čekovat jestli neni něco toho typu i v ne atomickejch
	vrcholech/typech...
\end{todo}

@@

\begin{todo}
& EXAMPLES of tree representations of \lterms  
& TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
		Things about $\beta\eta^{-1}$ normal form, etc.   
& DESCRIBE Inhabitation Machine... 
\end{todo}

@@

\begin{todo}
& Dřív se to menovalo 
"Our approach to \textit{solving} Inhabitation tree"
tak tu napsat pěknej uvod o tom, že naši jedinci jsou lambda
termy a že je  budeme generovat pomocí 
řešení inhabitačních stromu.
\end{todo}

@@

\textit{Classical-GP-like}

\begin{todo}
!!!
\end{todo}

@@

\begin{todo}
& víc to tu rozepsat hlavně zmínit heuristiku 
a že stačí aby byla optimistická atd..
\end{todo}

@@

\begin{todo}
& čeknout jak je to v kódu jestli ten A* co tam je čekuje ty
 nódy co už navštívil při přidávání - VŽDYŤ díky tomu žeje to tree-search 
 (vs. graph search) to neni potřeba - pořádně se podívat do AIAMA jakj
 je terminologie a napsat to sem.
\end{todo}

@@

\begin{todo}
 & Discuss that it is OK to expand only the DFS-found node. 
The order of unfinished nodes is irrelevant since generation of each branch is 
independent on generating other branches. Mužem na to koukat jako na takovou tu redukci simetrií, najít to v AIAMA a pořádně popsat.

& Odiskutovat nějak tohle:
přirozeně chceme při produkci velkýho množství termu obhospodařovávat nějaký
velký množství rozdělanejch termu v nějaký struktuře a postuipně je dodělávat.
Proto přichází jako tahle struktura na mysl nějaká prioritní fronta,
otázka je kde vzít ty priority a A* je jednoduchá ale přitom optimum hledající 
odpověď na tuhle otázku, dostatečně volná skrz tu heuristiku.

& určitě někde zmínit ospravedlnění proč si napsat dokazovač sám
a né požít nějakej už hotovej - de o to, že dokazovače jsou delaný na to aby
udělali jeden dukaz typicky složitýho tvrzení, my sme v situaci kdy
chceme co nejvíc (a co možná co nejvíc rozmanitejch) dukazu jednoduchýho problému.
\end{todo}

@@

\begin{todo}
 & rozebrat to, ukázat že je optimistická atd...
\end{todo}

@@

\begin{todo}
& Prostě budem křížit jen podstromy stejnýho typu
& ALE! Hlavní problém jsou proměnný
& ve stručnosti kouknout na dvě hlavní metody
  && křížit prostě lambda stromy a vzniklý problémy řešit online 
  && zbavit se proměnejch 
& říct že nakonec sme přilnuli k zbavit se proměnejch
\end{todo}

@@

\subsection{Abstraction elimination}
\label{toSKI}
%... Conversion to SKI combinators

\textit{Abstraction elimination} is a process of transforming 
an arbitrary \lterm into \lterm witch contains no lambda abstractions
and no bound variables.
The newly produced \lterm may contain function applications, 
free symbols from former \lterm and some new symbols standing for 
combinators $\Scomb$, $\Kcomb$ and $\Icomb$. \\

Those combinators are defined as:
\begin{align*}
\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
\Kcomb &= \lam{x\,y}{x} \\
\Icomb &= \lam{x}{x} 
\end{align*}

\newcommand{\Ae}{\mathop{\mathrm{\AE}}}

\Lets describe transformation $\Ae$ performing this process.
\begin{align*}
\Ae[x]           &= x &\\[0.4em]
\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
\Ae[\lam{x}{x}]  &= \Icomb &\\
\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
&\textbf{if } x \in \FV(M)\\
\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
&\textbf{if } x \in \FV(M) \vee x \in \FV(N)\\
\end{align*}

\begin{todo}
& přidat (ale i do kódu) B a C
& říct že my musíme řešit i typy 
& říct o složitosti , a že je to docela
disadventič vzhledem k bloatu
& citovat Simon Peyton Jones implementaition of FP languages
jako source of further reading on this subject.
\end{todo}

@@

\begin{todo}
& říct že na tu sme se moc nesoustředili ze dvou duvodu
  ktery to ospravedlňujou 
&& koza taky
&& ta s vygereovanim podstromu je v zásadě použití
   algoritmu na generovaní jedince a tak nám přišlo
   lepší pořádně se soustředit na generování jedince a
   začít do toho mutaci tahat až bude generování pořádně
   prozkoumáno.
\end{todo}

@@

\chapter{Comparison with others}
